---
title: "Regression and Classification Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(123)

# UCI ML datasets and many simulated datasets available
require(mlbench)
path=getwd()

# no of records
n = 100
# no of feature
p = 10

giveColNames <- function(p)
{
  cnames <- c('target',paste(rep("feature_",p),formatC(1:p,width=floor(1+log10(p)),flag=0,format="d"),sep=""))
}

writeData <- function(X,y,fname)
{
  p <- ncol(X)
  cnames <- c('target',paste(rep("feature_",p),formatC(1:p,width=floor(1+log10(p)),flag=0,format="d"),sep=""))
  # regression problem
  df <- as.data.frame(cbind(y,X))
  colnames(df) <- cnames
  write.csv(df,paste(fname,".regr.csv",sep=""),row.names = F)
  
  yhat <- y-median(y,na.rm = T)
  prob <- 1/(1+exp(-yhat))
  y <-rbinom(n,1,prob)
  df$target <- y
  write.csv(df,paste(fname,".class.csv",sep=""),row.names = F)
}
```


## 1. Single Feature

**Data**

Generate `r n` records with 1 features

**comments**

Should be straight forward

```{r 1}
set.seed(1)
X <- matrix(rnorm(n*p),ncol=1)
beta <- matrix(rnorm(1,0,10),ncol=1)
y <- X%*%beta
writeData(X,y,"Ex01")
```




## 2. Multiple Features

#### 2a. independent features

**Data**

Generate `r n` records with `r p` indepdent features.

**Comments**

* last two features are not important (with lasso, it should produce exact zero)

```{r 2a}
set.seed(2)
X <- matrix(rnorm(n*p),ncol=p)
beta <- matrix(rnorm(p,0,10),ncol=1); beta[c(p-1,p)] <- 0
y <- X%*%beta
writeData(X,y,"Ex02a")
```

#### 2b correlated features

**Data**

Generate `r n` records with `r p` correlated features.

**Comments**

* last two features are not important (with lasso, it should produce exact zero)
* gradient descent would be unstable
* variable selection is not consistent

```{r 2b}
set.seed(2)
p=10
n=100
X <- matrix(rnorm(n*p),ncol=p)
X[,3] <- 0.9*X[,1]
X[,p] <- 0.9*X[,1] -0.5*X[,3]  
beta <- matrix(rnorm(p,0,10),ncol=1); beta[c(p-1,p)] <- 0
y <- X%*%beta
writeData(X,y,"Ex02b")
```


#### 2c features of different scale

**Data**

Generate `r n` records with `r p` indepdent features. Each feature is on a different scale and different mean

**Comments**

* last two features are not important (with lasso, it should produce exact zero)
* gradient descent would be unstable


```{r 2c}
set.seed(2)
p=10
n=100
X <- matrix(rnorm(n*p),ncol=p)
X <- scale(X); X<- scale(X,center=rnorm(p,0,10),scale=abs(0.1+rnorm(p,0.5,15))) 
X[,2] <- rnorm(n,1,0.01)
beta <- matrix(rnorm(p,0,10),ncol=1); beta[c(p-1,p)] <- 0
y <- X%*%beta
writeData(X,y,"Ex02c")
```


#### 2d correlated features and with different scale

**Data**

Generate `r n` records with `r p` correlated features. Each feature is on a different scale and different mean

**Comments**

* last two features are not important (with lasso, it should produce exact zero)
* gradient descent would be unstable
* variable selection is not consistent

```{r 2d}
set.seed(2)
p=10
n=100
X <- matrix(rnorm(n*p),ncol=p)
X <- scale(X); X<- scale(X,center=rnorm(p,0,10),scale=abs(0.1+rnorm(p,0.5,15))) 
X[,2] <- rnorm(n,1,0.01)
X[,3] <- 0.9*X[,1]
X[,p] <- 0.9*X[,1] -0.5*X[,3]  
beta <- matrix(rnorm(p,0,10),ncol=1); beta[c(p-1,p)] <- 0
y <- X%*%beta
writeData(X,y,"Ex02d")
```


#### 2e MLR correlated features and with different scale, mising data and outliers.

**Data**

Generate `r n` records with `r p` correlated features. Each feature is on a different scale and different mean

**Comments**

* last two features are not important (with lasso, it should produce exact zero)
* gradient descent would be unstable
* variable selection is not consistent
* regression/classfication are noisy


```{r 2e}
set.seed(2)
p=10
n=100
X <- matrix(rnorm(n*p),ncol=p)
X <- scale(X); X<- scale(X,center=rnorm(p,0,10),scale=abs(0.1+rnorm(p,0.5,15))) 
X[,2] <- rnorm(n,1,0.01)
X[,3] <- 0.9*X[,1]
X[,p] <- 0.9*X[,1] -0.5*X[,3]

# plant missing data
xna.row <- sample(n,5,replace=FALSE)
xna.col <- sample(p,5,replace=TRUE)
X[cbind(xna.row,xna.col)] <- NA

#plan outlier
xna.row <- sample(n,2,replace=FALSE)
xna.col <- sample(p,2,replace=TRUE)
X[cbind(xna.row,xna.col)] <- 1e10
xna.row <- sample(n,2,replace=FALSE)
xna.col <- sample(p,2,replace=TRUE)
X[cbind(xna.row,xna.col)] <- -1e10
beta <- matrix(rnorm(p,0,10),ncol=1); beta[c(p-1,p)] <- 0
y <- X%*%beta
# only in target
yna <- sample(n,2,replace=FALSE)
y[yna] <- NA

writeData(X,y,"Ex02e")
```


